---
title: "Clinical Variant Analysis Report"
author: "Hussein Moussa"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(ggplot2)
library(dplyr)
```

## Overview

This report summarizes the analysis of genetic variants from the 
1000 Genomes Project, filtered by read depth and annotated 
using the ClinVar database.
```{r load_data}
df <- read.csv("annotated_variants.csv")

df$ClinVar_Significance <- case_when(
  grepl("athogenic", df$ClinVar_Significance) ~ "Pathogenic",
  grepl("enign",     df$ClinVar_Significance) ~ "Benign",
  grepl("ncertain",  df$ClinVar_Significance) ~ "Uncertain",
  TRUE                                         ~ "Not in ClinVar"
)

total     <- nrow(df)
pathogenic <- sum(df$ClinVar_Significance == "Pathogenic")
benign     <- sum(df$ClinVar_Significance == "Benign")
uncertain  <- sum(df$ClinVar_Significance == "Uncertain")
```

## Summary Statistics

| Metric | Count |
|--------|-------|
| Total Variants Analyzed | `r total` |
| Pathogenic | `r pathogenic` |
| Benign | `r benign` |
| Uncertain Significance | `r uncertain` |
| Not in ClinVar | `r total - pathogenic - benign - uncertain` |

## Variant Classification
```{r plot1, fig.width=7, fig.height=5}
ggplot(df, aes(x = ClinVar_Significance, fill = ClinVar_Significance)) +
  geom_bar() +
  scale_fill_manual(values = c(
    "Pathogenic"     = "#e74c3c",
    "Benign"         = "#2ecc71",
    "Uncertain"      = "#f39c12",
    "Not in ClinVar" = "#95a5a6"
  )) +
  labs(
    title = "Variant Classification by ClinVar Significance",
    x     = "Clinical Significance",
    y     = "Number of Variants"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Read Depth by Clinical Significance
```{r plot2, fig.width=7, fig.height=5}
ggplot(df, aes(x = ClinVar_Significance, y = DP, fill = ClinVar_Significance)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c(
    "Pathogenic"     = "#e74c3c",
    "Benign"         = "#2ecc71",
    "Uncertain"      = "#f39c12",
    "Not in ClinVar" = "#95a5a6"
  )) +
  labs(
    title = "Read Depth by Clinical Significance",
    x     = "Clinical Significance",
    y     = "Read Depth (DP)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

## Pathogenic Variants
```{r pathogenic_table}
pathogenic_df <- df[df$ClinVar_Significance == "Pathogenic", 
                    c("CHROM", "POS", "ID", "REF", "ALT", "DP")]

knitr::kable(pathogenic_df, 
             caption = "Confirmed Pathogenic Variants",
             row.names = FALSE)
```

## Conclusion

This pipeline successfully parsed, filtered, and annotated a sample of 100 genetic 
variants from the 1000 Genomes Project. Pathogenic variants were 
identified and confirmed against the ClinVar database. This workflow 
mirrors standard practices used in clinical genomics diagnostic labs.